risk-perception

### 2019-10-15 risk-percption.Rmd
plots

calc_yield <- function(wheat_prop, risk) {
  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
  barley + wheat
}

util_wet * (1 - risk)

calc_yield(.92, .5)

calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)

calc_yield(1, .2)

plot(calc_yield((0:10)/10, .2))

expand_grid(wheat_prop = (0:10)/10,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

expand_grid(wheat_prop = (0:10)/1,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(risk, util, color = type)) + geom_line() + facet_wrap(~wheat_prop)

@@

plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)
```

@@

tibble(rainfall = (0:200) / 100,
       barley = calc_yield(rainfall, 'barley'),
       wheat = calc_yield(rainfall, 'wheat')) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + 
  geom_line() +
  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
  geom_hline(yintercept = yield_max_barley * kcal_barley)
```

>>>>

calc_payoff <- function(wheat_prop, risk) {
  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
  barley + wheat
}

tibble(wheat_prop)
```

calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)
```

So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
```{r}
test_dat <- expand_grid(wheat_prop = (0:10)/10,
       risk = seq(0, 1, 0.2)) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk)+ barley_dry * risk),
         payoff = calc_payoff(wheat_prop, risk))

test_dat %>%
  group_by(risk) %>%
  filter(payoff == max(payoff)) %>%
  ggplot(aes(risk, wheat_prop)) + geom_point()

test_dat %>%
  select(-(util_dry:util_wet)) %>%
  gather(type, util, util_wheat:util_barley) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
test_dat %>%
  select(-(util_wheat:util_barley)) %>%
  gather(type, util, util_dry:util_wet) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

test_dat %>%
  ggplot(aes(wheat_prop, util_wet + util_dry)) + geom_point() + facet_wrap(~risk)

@@

plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)

tibble()

@@

tibble(rainfall = (0:200) / 100,
       barley = calc_yield(rainfall, 'barley'),
       wheat = calc_yield(rainfall, 'wheat')) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + 
  geom_line() +
  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
  geom_hline(yintercept = yield_max_barley * kcal_barley) +
  geom_hline(yintercept = 1016525* 5, color ='red') +
  geom_vline(xintercept = 0.3)
```

How do we define the "drought year" cutoff? Below 300mm there isn't enough wheat growing in a hectare to feed a family of 5, which seems appropriate.
```{r}
test <- tibble(time = 1:100,
       rain = rnorm(100, mean = 1, sd = 1) %>% pmax(0),
       yield = calc_yield(rain))

ggplot(test, aes(time, yield)) + geom_point() + geom_hline(yintercept = 1016525)

test %>%
  mutate(drought = if_else(yield < 5e+6, TRUE, FALSE))%>%
  filter(drought == TRUE)
```
